The extended light profiles of the Sculptor and Ursa Minor dwarf galaxies: Tides or nature?

### 2025-08-31 thesis/abstract.tex
thesis julio ch1-3 edits

The satellite galaxies of the Milky Way (MW), with the exception of the
Magellanic clouds, are all dwarf spheroidals (dSphs)---gas-free,
non-rotating stellar systems who's gravity is dominated by dark matter.
The light profiles of dSphs are usually well-approximated by an
exponential law, with a sharp outer cutoff. Yet systems like the
Sculptor (Scl) and Ursa Minor (UMi) dSphs have been found to host member
stars far from the main body, out to \textasciitilde10 effective radii.
The origin of these extended profiles is not clear, and may reflect
either the influence of Galactic tides or instead an innate feature. In
this thesis, I review the excestence of an extended light profiles in
Scl and UMi, validating the Bayesian membership catalogue of
\citet{jensen+2024}. To evaluate if tides indeed can explain these
density features, I conduct idealized N-body simulation of both galaxies
in the tidal field of the Milky Way. I find that neigher dwarf has been
subjected to tidal forces strong enough to create the observed extended
light profiles from initially exponential profiles. One complication is
that Sculptor's orbit is strongly influenced by the presence of the
Large Magellanic Cloud (LMC). Because the LMC's mass is poorly
constrained, Sculptor's long term orbital history is highly uncertain.
Despite this uncertainty, our models still suggest that the combined
tides of the LMC and MW are still unable to explain Scl's outer profile
over the past 5 Gyr. I conclude that the extended light profiles of Scl
and UMi are innate, likely reflecting their formation histories.
Mergers, accretion events, or complex, multi-component star formation
episodes are possible explanations for such an extended outer profile.

>>>>

The satellite galaxies of the Milky Way (MW), with the exception of the
Magellanic clouds, are all dwarf spheroidals (dSphs)---gas-free,
non-rotating stellar systems whos gravity is dominated by dark matter.
The light profiles of dSphs are usually well-approximated by an
exponential law, with a sharp outer cutoff. Yet systems like the
Sculptor (Scl) and Ursa Minor (UMi) dSphs have been found to host member
stars far from the main body, out to \(\sim10\) effective radii. The
origin of these extended profiles is not clear, and may reflect either
the influence of Galactic tides or instead an innate feature. In this
thesis, I review the excestence of an extended light profiles in Scl and
UMi, validating the Bayesian membership catalogue of
\citet{jensen+2024}. To evaluate if tides indeed can explain these
density features, I conduct idealized N-body simulation of both galaxies
in the tidal field of the Milky Way. I find that neigher dwarf has been
subjected to tidal forces strong enough to create the observed extended
light profiles from initially exponential profiles. One complication is
that Sculptor's orbit is strongly influenced by the presence of the
Large Magellanic Cloud (LMC). Because the LMC's mass is poorly
constrained, Sculptor's long term orbital history is highly uncertain.
Despite this uncertainty, our models still suggest that the combined
tides of the LMC and MW are still unable to explain Scl's outer profile
over the past 5 Gyr. I conclude that the extended light profiles of Scl
and UMi are innate, likely reflecting their formation histories.
Mergers, accretion events, or complex, multi-component star formation
episodes are possible explanations for such an extended outer profile.

### 2025-08-31 thesis/data.tex
thesis julio ch1-3 edits

As discussed in Section~\ref{sec:exponential_profiles}, the projected
luminosity/stellar mass density profile of dwarf galaxies is generally
well-described by an exponential law Eq.~\ref{eq:exponential_profiles}.
A prototypical example is the Fornax dSph --- well-fit by an exponential
over 4.5 decades in surface density. On the other hand, dSph like
Sculptor or Ursa Minor have profiles with deviate significantly from an
exponential profile fitted to the inner regions, with a clear excess of
stars/light in the outer regions. In this chapter, we critically review
the density profile of MW classical dwarfs. This thesis concerns the
origin of this excess, with a tidal interpretation explored in
subsequent chapters.

\textbf{question: how much about density profiles in this
paragraph\ldots{}}

@@

For our fiducial sample, we adopt a minimum probability of
\(P_{\rm sat} = 0.2\). Most stars have \(P_{\rm  sat}\) values which are
nearly 0 or 1, so the exact choice of probability threshold has little
effect on the resulting sample. Even at our relatively generous
probability threshold of 0.2, the purity remains high when validated
against spectroscopic line-of-sight (LOS) velocities
(\textasciitilde90\%, J+24). We note, however, that these purity
estimates may be biased. Stars with LOS velocities are typically
brighter and, as a result, have more precise \emph{Gaia} measurements.
However, we find that our conclusions are unchanged when limiting
samples to only the brightest stars. Altogether, the J+24 method
provides a high-quality, low-contamination sample of dwarf galaxy
candidate member stars, which we will now investigate.

@@

We analyze stellar distributions in the tangent plane, considering as
well the projected shape of a galaxy. The tangent plane coordinates
\(\xi\) and \(\eta\) are offsets in RA and declination as measured on
the plane tangent to the galaxy centre. To account for the elliptical
shape of the galaxy, we use \(R_{\rm ell}\), which we define as the
circularized elliptical radius, \begin{equation}{
R_{\rm ell}^2 = a\,b\,({\xi'}^2 / a^2 + {\eta'}^2 / b^2),
}\end{equation} where \(\xi'\) and \(\eta'\) are the tangent plane
coordinate rotated to align with a dwarf galaxy's major and minor axis,
and \(a\) and \(b\) are the semi-major and semi-minor axis of the
galaxy.

We illustrate the likelihood-based membership selection criteria through
a progressive tightening of criteria. First, we consider a
minimally-refined sample, \textbf{all}, only excluding stars with poor
astrometry, unreliable photometry, or inconsistent parallaxes. Next,
incorporating CMD and PM information, we use the \textbf{CMD+PM} to test
a selection method agnostic to the spatial position. This sample
includes stars where the CMD and PM combined likelihood favours
satellite membership:
\begin{equation}\protect\phantomsection\label{eq:sel_cmd_pm}{
{\cal L}_{\rm CMD,\ sat}\ {\cal L}_{\rm PM,\ sat} > {\cal L}_{\rm CMD,\ bg}\ {\cal L}_{\rm PM,\ bg}.
}\end{equation} Next, our \textbf{fiducial sample},
\(P_{\rm sat} > 0.2\), includes a spatial likelihood as described in
J+24. Finally, the \textbf{RV members} sample adds line-of-sight
velocity information (see Appendix for details). While the latter is the
best motivated membership sample, we do not use this sample for stellar
density analysis due to its incompleteness and complex selection
function.

Figs.~\ref{fig:scl_selection}, \ref{fig:umi_selection}, \ref{fig:fornax_selection}
shows our fiducial sample for each galaxy, and the impact of alternate
samples, in tangent-plane coordinates, the CMD, and in proper-motion
space. The \emph{all} sample extends uniformly across the tangent plane,
but includes a substantial background population. The \emph{CMD+PM}
sample has a much lower background density in the tangent plane,
revealing the satellite more clearly. The fiducial sample appears
similar to the \emph{CMD+PM} sample in the CMD and PM planes, but
excludes improbably distant stars. The \emph{RV member} sample also
traces out similar distributions in CMD and PM space as the fiducial
sample, with less dispersion, likely reflecting the brighter magnitudes
of stars with spectroscopic follow-up. Each component plays an important
role in ensuring a high-quality membership criterion.

Based on the fiducial sample's distribution in
Figs.~\ref{fig:scl_selection}, \ref{fig:umi_selection}, \ref{fig:fornax_selection},
Ursa Minor stands out because of its higher ellipticity. Sculptor and
Ursa Minor show a population of members past \(6R_h\) in radius, but
Fornax does not, even though Fornax has many more stars. Fornax also has
a bluer CMD, indicative of more recent star formation. Otherwise, there
are no standout morphological features between these galaxies.

One limitation of the J+24 method is the assumption of a specific
density profile for the dwarf galaxy (a one- or two-component
exponential), which may impact the membership probability of distant
stars. However, there are no clear extensions or overdensities in the
\emph{CMD+PM} selected sample, which does not include a spatial
likelihood. If even fainter, more extended features exist, they are
apparently not detectable with \emph{Gaia}.

To illustrate the extent of Sculptor and Ursa Minor, we highlight the
most distant, spectroscopically confirmed members with red-outlined
stars in Figs.~\ref{fig:scl_selection}, \ref{fig:umi_selection}. In
particular, \citet{sestito+2023a}; \citet{sestito+2023b} targeted the
most distant stars using the J+24 candidate membership list. These stars
illustrate an extended stellar distribution for each galaxy. For an
exponential profile, 99.95\% of stars fall within a radius of \(6R_h\),
so the mere presence of very distant member stars provides direct
evidence for deviations from a simple, exponential profile.

\begin{figure}
\centering
\includegraphics[width=0.7\linewidth,height=\textheight,keepaspectratio]{figures/scl_selection.png}
\caption[Sculptor sample selection]{The distributions of various samples
of \emph{Gaia} stars for Sculptor. We plot light grey points for field
stars (with consistent parallaxes and reliable photometry and
astrometry), turquoise points for \emph{CMD+PM} selected stars
(Eq.~\ref{eq:sel_cmd_pm}), blue \texttt{x}s for the fiducial sample
(\(P_{\rm sat} > 0.2\)), and indigo diamonds for the RV confirmed
members. We mark the two far-outlier stars from \citet{sestito+2023a}
with rust-outlined indigo stars. \textbf{Top:} Tangent plane
\(\xi, \eta\). The orange ellipse represents 3 half-light radii.
\textbf{Bottom left:} Colour magnitude diagram in Gaia \(G\) magnitude
versus \(G_{\rm BP} - G_{\rm RP}\) colour. We plot a Padova \(12\,\)Gyr,
\({\rm [Fe/H] }=-1.68\) isochone in orange. The black bar in the top
left represents the median colour error. \textbf{Bottom right:} Proper
motion in declination \(\mu_\delta\) vs RA \(\mu_{\alpha*}\)
(corrected). The orange point marks the systemic \citet{MV2020b} proper
motion. The black cross represents the median proper motion error.
\textbf{TODO: reduce png sizes}}\label{fig:scl_selection}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.7\linewidth,height=\textheight,keepaspectratio]{figures/umi_selection.png}
\caption[Ursa Minor sample selection]{Similar to
Fig.~\ref{fig:scl_selection} except for Ursa Minor. We outline RV
members outside a radius of \(6R_h\) with red stars (from
\citet{sestito+2023b}, \citet{pace+2020} and \citet{spencer+2018}).
\textbf{TODO: plot 6Rh ellipse\ldots{}}}\label{fig:umi_selection}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.7\linewidth,height=\textheight,keepaspectratio]{figures/fornax_selection.png}
\caption[Fornax sample selection]{Similar to
Fig.~\ref{fig:scl_selection} except for Fornax. RV members are from
\citet{WMO2009}. Fornax does not show the same extended outer halo of
probable members as Sculptor or Ursa Minor despite having many more
stars.}\label{fig:fornax_selection}
\end{figure}

@@

Fig.~\ref{fig:scl_observed_profiles} show the derived density profiles
for Sculptor, Ursa Minor, and Fornax. We calculate density profiles for
different selections of stars from above: \emph{all} stars, CMD+PM only,
and the fiducial sample. In each case, all samples are the same towards
the inner regions of the satellite. However, the density profile
\emph{quality stars} plateaus at the total background in the field at
radii of 30-60 arcminutes, depending on the galaxy. By restricting
stellar membership with CMD + PM, the background is reduced by 1-2 dex.
The CMD+PM background plateau represents the density of background stars
which could be mistaken as members because of their coincident colours
and proper motions. Finally, the background (BG)-subtracted profile
results from subtracting the apparent background in the \emph{all}
profile.

\textbf{Describe R\_break, etc. in Fig.2.4} Note that, in all three
galaxies, the data does not extend out to the break radius.

While the density profiles of all samples agrees in the inner regions,
they begin to deviate outside \(6 R_h\). Outside this radius, the
fiducial profile extends 1--2 magnitudes below the CMD+PM selection
background. Because the fiducial profile in this region depends
critically on the spatial likelihood, its shape is vulnerable to the
assumed density profile (an exponential in J+24). We explore this in
more detail in Appendix {[}@-sec:density\_extra{]}.

Nearto UMi, there is a small, likely unassociated, ultrafaint star
cluster, Muñoz 1 \citep{munoz+2012}. The cluster is at a relative
position of \((\xi, \eta) \approx(-42, -15)\) arcminutes, corresponding
to an elliptical radius of 36 arcminutes. This cluster does not have a
bright RGB, so has few stars brighter than a magnitude \(G=22\). The
cluster has little effect on the elliptically-averaged density profile.
\textbf{Mark Muñoz 1}

Fig.~\ref{fig:classical_dwarfs_densities} compares the fiducial density
profiles of Sculptor, Ursa Minor, Fornax, and other classical dwarf
galaxies. Of the classicals, we exclude Antlia II due to the extremely
high background, and Sagittarius which was not included in J+24. The
density profiles are scaled to match at the half-light radius from
\citet{munoz+2018}. All of the classical dwarfs appear to be well
described by an exponential profile in the inner regions. In the outer
regions, however, Sculptor and Ursa Minor deviate, and show a clear
outer excess over an exponential law (solid black line). These galaxies
are better fit by a Plummer law (dashed black line). The deviation from
an exponential grows outwards, and at \(\sim 8 R_h\), may reach 2 orders
of magnitude. This excess is remarkable because the spatial likelihood
used to derive the fiducial sample probabilities is the same (a single
component exponential) in all cases. The remainder of this thesis will
be devoted to assessing whether the outer excess shown by Scl and UMi
are due to Galactic tides.

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{figures/scl_umi_fnx_density_methods.pdf}}
\caption[Sculptor density profiles]{The density profile of Sculptor,
Ursa Minor, and Fornax for different selection criteria, ploted as log
surface dencity versus log elliptical radius. \emph{CMD+PM} select stars
more likely to be members according to CMD and PM only, \emph{all}
selects any high quality star, and \emph{BG subtracted} is the
background-subtracted density derived from \emph{all} stars. We mark the
half-light radius (vertical dashed line) and the break radius (black
arrow, Eq.~\ref{eq:r_break}). \textbf{rename
bg-subtracted}}\label{fig:scl_observed_profiles}
\end{figure}

>>>>

As discussed in Section~\ref{sec:exponential_profiles}, the projected
luminosity/stellar mass density profile of dwarf galaxies is generally
well-described by an exponential law (Eq.~\ref{eq:exponential_law}). A
prototypical example is the Fornax dSph --- well-fit by an exponential
over 4.5 decades in surface density. On the other hand, dSph like
Sculptor or Ursa Minor have profiles with deviate significantly from an
exponential profile fitted to the inner regions, with a clear excess of
stars/light in the outer regions. In this chapter, we critically review
the density profile of MW classical dwarfs. This thesis concerns the
origin of this excess, with a tidal interpretation explored in
subsequent chapters.

@@

For our fiducial sample, we adopt a minimum probability of
\(P_{\rm sat} = 0.2\). We use the \(P_{\rm sat}\) values from the
elliptical 2-component runs if a galaxy showed evidence for an outer
component, the 1-component run otherwise. Most stars have
\(P_{\rm  sat}\) values which are nearly 0 or 1, so the exact choice of
probability threshold has little effect on the resulting sample. Even at
our relatively generous probability threshold of 0.2, the purity remains
high when validated against spectroscopic line-of-sight (LOS) velocities
(\textasciitilde90\%, J+24). We note, however, that these purity
estimates may be biased. Stars with LOS velocities are typically
brighter and, as a result, have more precise \emph{Gaia} measurements.
However, we find that our conclusions are unchanged when limiting
samples to only the brightest stars. Altogether, the J+24 method
provides a high-quality, low-contamination sample of dwarf galaxy
candidate member stars, which we will now investigate.

@@

We analyze stellar distributions in the tangent plane, considering as
well the projected shape of a galaxy. The tangent plane coordinates
\(\xi\) and \(\eta\) are offsets in RA and declination as measured on
the plane tangent to the galaxy centre. To account for the elliptical
shape of the galaxy, we use \(R_{\rm ell}\), which we define as the
circularized elliptical radius, \begin{equation}{
R_{\rm ell}^2 = a\,b\,\left(\frac{{\xi'}^2}{a^2} + \frac{{\eta'}^2}{b^2} \right),
}\end{equation} where \(\xi'\) and \(\eta'\) are the tangent plane
coordinate rotated to align with a dwarf galaxy's major and minor axis,
and \(a\) and \(b\) are the semi-major and semi-minor axis of the
galaxy.

We illustrate the likelihood-based membership selection criteria through
a progressive tightening of criteria. First, we consider a
minimally-refined sample, \textbf{all}, only excluding stars with poor
astrometry, unreliable photometry, or inconsistent parallaxes. Next,
incorporating CMD and PM information, we use the \textbf{CMD+PM} to test
a selection method agnostic to the spatial position. This sample
includes stars where the CMD and PM combined likelihood favours
satellite membership:
\begin{equation}\protect\phantomsection\label{eq:sel_cmd_pm}{
{\cal L}_{\rm CMD,\ sat}\ {\cal L}_{\rm PM,\ sat} > {\cal L}_{\rm CMD,\ bg}\ {\cal L}_{\rm PM,\ bg}.
}\end{equation} Next, our \textbf{fiducial} sample,
\(P_{\rm sat} > 0.2\), includes a spatial likelihood as described in
J+24. Finally, the \textbf{RV members} (for radial velocity) sample adds
line-of-sight velocity information (see Appendix \ref{sec:rv_obs} for
details). While the latter is the best motivated membership sample, we
do not use this sample for stellar density analysis due to its
incompleteness and complex selection function.

Figs.~\ref{fig:scl_selection}, \ref{fig:umi_selection}, \ref{fig:fornax_selection}
shows our fiducial sample for each galaxy, and the impact of alternate
samples, in tangent-plane coordinates, the CMD, and proper-motion space.
The ``all'' sample extends uniformly across the tangent plane, but
includes a substantial background population. The ``CMD+PM'' sample has
a much lower background density in the tangent plane, revealing the
satellite more clearly. The fiducial sample appears similar to the
``CMD+PM'' sample in the CMD and PM planes, but excludes improbably
distant stars. The ``RV member'' sample also traces out similar
distributions in CMD and PM space as the fiducial sample, with less
dispersion, likely reflecting the brighter magnitudes of stars with
spectroscopic follow-up. Each selection criteria---spatial, CMD, and
PM---contributes towards a high-quality membership criterion.

Based on the fiducial sample's distribution in
Figs.~\ref{fig:scl_selection}, \ref{fig:umi_selection}, \ref{fig:fornax_selection},
Ursa Minor stands out because of its higher ellipticity. Sculptor and
Ursa Minor show a population of members past \(6R_h\) in radius, but
Fornax does not, even though Fornax has many more stars. Fornax also has
a bluer CMD, indicative of more recent star formation. Otherwise, there
are no major morphological differences between these galaxies.

One limitation of the J+24 method is the assumption of a specific
density profile for the dwarf galaxy (a one- or two-component
exponential), which may impact the membership probability of distant
stars. However, there are no clear extensions or overdensities in the
``CMD+PM'' selected sample, which does not include a spatial likelihood.
If even fainter, more extended features exist, they are apparently not
detectable with \emph{Gaia}.

To illustrate the extent of Sculptor and Ursa Minor, we highlight
distant, spectroscopically confirmed members with red-outlined stars in
Figs.~\ref{fig:scl_selection}, \ref{fig:umi_selection}. In particular,
\citet{sestito+2023a}; \citet{sestito+2023b} targeted distant, bright
stars from the J+24 candidate membership list. The most distant stars
confirmed in their work lie at radii \(\gtrsim 10 R_h\) from the centres
of each dwarf. For an exponential profile, 99.95\% of stars fall within
a radius of \(6R_h\), so the presence of very distant member stars
provides evidence for deviations from an exponential profile.

\begin{figure}
\centering
\includegraphics[width=0.7\linewidth,height=\textheight,keepaspectratio]{figures/scl_selection.png}
\caption[Sculptor sample selection]{The distributions of various samples
of \emph{Gaia} stars for Sculptor. We plot light grey points for ``all''
field stars (with consistent parallaxes and reliable photometry and
astrometry), turquoise points for ``CMD+PM'' selected stars
(Eq.~\ref{eq:sel_cmd_pm}), blue squares for the ``fiducial'' sample
(with membership probability \(P_{\rm sat} > 0.2\)), and indigo diamonds
for the ``RV members'' sample. We mark the two far-outlier stars from
\citet{sestito+2023a} with rust-outlined indigo stars. \textbf{Top:}
Tangent plane \(\xi, \eta\). The orange ellipses represents 3 and 6
half-light radii. \textbf{Bottom left:} Colour magnitude diagram in Gaia
\(G\) magnitude versus \(G_{\rm BP} - G_{\rm RP}\) colour. We plot a
Padova \(12\,\)Gyr, \({\rm [Fe/H] }=-1.68\) isochone in orange. The
black bar in the top left represents the median colour error.
\textbf{Bottom right:} Proper motion in declination \(\mu_\delta\) vs RA
\(\mu_{\alpha*}\) (corrected). The orange point marks the systemic
\citet{MV2020b} proper motion. The black cross represents the median
proper motion error.}\label{fig:scl_selection}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.7\linewidth,height=\textheight,keepaspectratio]{figures/umi_selection.png}
\caption[Ursa Minor sample selection]{Similar to
Fig.~\ref{fig:scl_selection} except for Ursa Minor. We outline
``velocity confirmed'' members outside a radius of \(6R_h\) with red
stars \citep[from][]{sestito+2023b, pace+2020, spencer+2018}. We also
mark the location of Muñoz 1 with a pink
circle.}\label{fig:umi_selection}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.7\linewidth,height=\textheight,keepaspectratio]{figures/fornax_selection.png}
\caption[Fornax sample selection]{Similar to
Fig.~\ref{fig:scl_selection} except for Fornax. RV measurements are from
\citet{WMO2009}. Fornax does not show the same extended outer halo of
probable members as Sculptor or Ursa Minor despite having many more
stars.}\label{fig:fornax_selection}
\end{figure}

@@

Fig.~\ref{fig:scl_observed_profiles} show the derived density profiles
for Sculptor, Ursa Minor, and Fornax. We calculate density profiles for
three different samples from above: ``all'', ``CMD+PM'', and the
``fiducial'' sample. In each case, all samples are the same towards the
inner regions of the satellite. However, the density profile from
``all'' plateaus at the total background in the field at radii of 30-60
arcminutes, depending on the galaxy. By restricting stellar membership
with CMD + PM, the background is reduced by 1-2 dex. The ``CMD+PM''
background plateau represents the density of background stars which
could be mistaken as members because of their coincident colours and
proper motions. Finally, the ``all -- background'' profile results from
subtracting the apparent background in the ``all'' profile.

While the density profiles of all samples agrees in the inner regions,
they begin to deviate outside \(6 R_h\). Outside this radius, the
fiducial profile extends 1--2 magnitudes below the CMD+PM selection
background. Because the fiducial profile in this region depends
critically on the spatial likelihood, its shape is vulnerable to the
assumed density profile (an exponential in J+24). We explore this in
more detail in Appendix \ref{sec:density_extra}.

Near to UMi, there is a small (\(R_h\sim 0.5'\)), likely unassociated,
ultrafaint star cluster, Muñoz 1 \citep{munoz+2012}. The cluster is at a
relative position of \((\xi, \eta) \approx(-42, -15)\) arcminutes,
corresponding to an elliptical radius of 37 arcminutes. This cluster
does not have a bright RGB, so has few stars brighter than a magnitude
\(G=22\). The cluster has little effect on the elliptically-averaged
density profile (see location on Fig.~\ref{fig:scl_observed_profiles}).

Fig.~\ref{fig:classical_dwarfs_densities} compares the fiducial density
profiles of Sculptor, Ursa Minor, Fornax, and other classical dwarf
galaxies. Of the classicals, we exclude Antlia II, due to the extremely
high background, and Sagittarius, which was not included in J+24. The
density profiles are scaled to match at the half-light radius from
\citet{munoz+2018}. All of the classical dwarfs appear to be well
described by an exponential profile in the inner regions. In the outer
regions, however, Sculptor and Ursa Minor deviate, and show a clear
outer excess over an exponential law (solid black line). These galaxies
are better fit by a Plummer law (dashed black line). The deviation from
an exponential grows outwards, and at \(\sim 8 R_h\), may reach 2 orders
of magnitude. The remainder of this thesis will be devoted to assessing
whether the outer excess shown by Scl and UMi are due to Galactic tides.

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{figures/scl_umi_fnx_density_methods.pdf}}
\caption[Sculptor density profiles]{The density profile of Sculptor,
Ursa Minor, and Fornax for different selection criteria, ploted as log
surface dencity versus log elliptical radius. The samples are: ``all''
selects any high quality and parallax-consistent star (grey
circles),``all--background'' subtracts a uniform background density from
the ``all'' profile (orange pentagons), ``CMD+PM'' select stars
according to CMD and PM only (teal triangles), and ``fiducial'' also
includes spatial information (blue squares). We mark the half-light
radius with a vertical dashed line and the background density with the
horizontal grey line. For Ursa Minor, we show the expected location of
Muñoz 1 stars as a horizontal bar (ranging from plus/minus 3 half-light
radii).}\label{fig:scl_observed_profiles}
\end{figure}
